Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-22 paper/paper.tex
Updates to the conclusions

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric, $T_C$. These plots confirm our previous observations about the
algorithms' performance, but they also allow us observe a more general property:
for a given budget, when the deadline is tightened the amount of work done
steeply decreases. Assigning a shorter deadline means that more VMs need to be
provisioned to complete the work. However, because of the dependencies in the 
workflows, many VMs will either be idle, or will be backfilled with tasks from 
low-priority workflows that never finish, resulting in reduced resource 
utilization. Similarly, the only way to reduce cost when adding VMs is to 
achieve a super-linear speedup, which is unlikely in practice. Therefore, if 
the deadline permits, and cost is important, parallelism should be kept to a 
minimum.

These observations are further confirmed when examining the effective computing
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increases in the effective computing cost. This increase can be 
significant. For example, a SIPHT ensemble with short deadlines and a budget
of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
minimum effective cost of \$1/hour, depending on the algorithm used. That means that out of the 
\$1,000 budget, \$80--\$200 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We observe that the selection of deadline and cost can be formulated as a
multi-criteria optimization problem, and that the curves presented in 
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the amount of work done, which directly results in minimization of effective 
computing cost metric for a given deadline. Therefore, the obtained results 
can be used to aid in resource allocation planning, when both cost and time 
criteria are given not as constraints but as objectives.

@@

Similar results were achieved for ensembles consisting of workflows of equal
size. A summary of these results is shown in Table~\ref{tab:num-dags-equal}.
Again, we can see that the static SPSS algorithm performed better for coarse-grained
workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
are smaller since the workflows have equal size, therefore rejecting a particular
workflow does not enable the algorithms to complete any other workflows.

Fig.~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
perspective. It helps answer the question: Given a deadline, how many workflows
can be finished when we increase the budget? In Fig.~\ref{fig:budgets}(a) the
growth is super-linear, which results from the Pareto-like distribution of 
workflow sizes in the ensemble (see Fig.~\ref{fig:ensemble-distribution}). 
For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
the growth is nearly linear.

@@

Another general observation is that although IaaS clouds provide elasticity
and the illusion of unlimited resources, which makes it possible to provision
more VMs to run workflows in parallel, it turns out that higher resource
utilization and better cost effectiveness can be achieved by limiting 
parallelism. This observation comes from the fact that 
starting and stopping VMs incurs overheads which can be amortized by running
each VM for as long as possible. This effect was confirmed in the case of 
the SPSS algorithm for Montage and CyberShake workflows with small budgets.
In these cases SPSS decided to provision many VMs in parallel, and shut them down early,
so that tasks early in the ensemble workflows could be completed by their 
deadlines. Similarly, a high degree of parallelization leads to lower 
resource utilization and increased cost. This result, visible in 
Fig~\ref{fig:cost}, leads to multi-criteria decision problems and related 
trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ensemble.

>>>>

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric, $T_C$. These plots confirm our previous observations about the
algorithms' performance, but they also allow us observe a more general property:
for a given budget, when the deadline is tightened the amount of work done
steeply decreases. This is partially due to the fact that fewer workflows can be
completed given a short deadline because their critical paths are simply too long
and they are squeezed out of early start times by high priority workflows.
In addition, assigning a shorter deadline means that more VMs need to be
provisioned to complete the work. However, because of the dependencies in the 
workflows, many of these VMs will either be idle, or will be backfilled with 
tasks from low-priority workflows that never finish, resulting, effectively, 
in reduced resource utilization and fewer completed workflows.

These observations are further confirmed when examining the effective computing
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increases in the effective computing cost. This increase can be 
significant. For example, a SIPHT ensemble with short deadlines and a budget
of \$1,000 has an effective computing cost $\sim$8--20\% higher than the 
minimum effective cost of \$1/hour, depending on the algorithm used. This means 
that out of the \$1,000 budget, \$80--\$200 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We observe that the selection of deadline and cost can be formulated as a
multi-criteria optimization problem, and that the curves presented in 
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the amount of work done, which is the inverse of minimizing the effective 
computing cost metric for a given deadline. Therefore, the obtained results 
can be used to aid in resource allocation planning, when both cost and time 
criteria are given not as constraints but as objectives.

@@

Similar results were achieved for ensembles consisting of workflows of equal
size. A summary of these results is shown in Table~\ref{tab:num-dags-equal}.
Again, we can see that the static SPSS algorithm performed better for coarse-grained
workflows (LIGO, Epigenomics, SIPHT). The differences between algorithm scores
are smaller since the workflows have equal size, therefore rejecting a particular
workflow does not make it any easier for the algorithms to complete any other 
workflow.

Fig.~\ref{fig:budgets} shows the same data for the SIPHT ensemble from another
perspective. It helps answer the question: Given a deadline, how many workflows
can be finished when we increase the budget? In Fig.~\ref{fig:budgets}(a) the
growth is super-linear, which results from the Pareto-like distribution of 
workflow sizes in the ensemble (see Fig.~\ref{fig:ensemble-distribution}). 
For ensembles of constant size workflows (see Fig.~\ref{fig:budgets}(b)) 
the growth is linear.

@@
